Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations

### 2021-03-25 papers/intro_paper/abstract.tex
minor changes to abstract and intro

  The no-hair conjecture in General Relativity (GR) states that the
  properties of an astrophysical Kerr black hole (BH) are completely described by its
  mass and spin angular momentum. As a consequence, the complex
  quasi-normal-mode (QNM) frequencies of a binary--black-hole (BBH)
  ringdown can be uniquely determined by the mass and spin of the
  remnant object. Conversely, measurement of the QNM frequencies could
  be an independent test of the no-hair conjecture. This paper extends to spinning BHs earlier work that proposed to
  test the no-hair conjecture by measuring the complex QNM
  frequencies of a BBH ringdown using parameterized inspiral-merger-ringdown waveforms in the effective-one-body formalism, 
thereby taking full advantage of the entire signal power and removing dependency on the
  predicted or estimated start time of the ringdown. Our method was used to analyse the
  properties of the merger remnants for four BBHs observed by
  LIGO-Virgo in the first half of their third observing (O3a) run in
  the latest LIGO-Virgo publication. After testing our method with GR and non-GR synthetic-signal injections in Gaussian noise, 
we analyse for the first time four other GW events: two BBHs from the first (O1) and second (O2) LIGO-Virgo 
  observing runs, and two BBHs from the O3a run. We then provide joint constraints with published results
  from the O3a run. In the most agnostic and conservative scenario where we combine the information from different
events using a hierarchical approach, we obtain, at $90\%$ credibility, that 
the fractional deviations in the frequency (damping time) of the dominant QNM are 
$\df{220}=0.02^{+0.09}_{-0.09}$ ($\dtau{220}=0.13^{+0.42}_{-0.40}$), respectively, an improvement of a factor of $\sim 4$ ($\sim 2$) 
over the results obtained with our model in the LIGO-Virgo publication. The single-event most-stringent constraint to date continues to be 
GW150914 for which we obtain $\df{220}=0.04^{+0.06}_{-0.04}$ and $\dtau{220}=0.09^{+0.18}_{-0.18}$.

>>>>

  The no-hair conjecture in General Relativity (GR) states that the
  properties of an astrophysical Kerr black hole (BH) are completely described by its
  mass and spin angular momentum. As a consequence, the complex
  quasi-normal-mode (QNM) frequencies of a binary--black-hole (BBH)
  ringdown can be uniquely determined by the mass and spin of the
  remnant object. Conversely, measurement of the QNM frequencies could
  be an independent test of the no-hair conjecture. This paper extends to spinning BHs earlier work that proposed to
  test the no-hair conjecture by measuring the complex QNM
  frequencies of a BBH ringdown using parameterized inspiral-merger-ringdown waveforms in the effective-one-body formalism, 
thereby taking full advantage of the entire signal power and removing dependency on the
  predicted or estimated start time of the ringdown. Our method was used to analyse the
  properties of the merger remnants for BBHs observed by
  LIGO-Virgo in the first half of their third observing (O3a) run in
  the latest LIGO-Virgo publication. After testing our method with GR and non-GR synthetic-signal injections in Gaussian noise, 
we analyse for the first time four other GW events: two BBHs from the first (O1) and second (O2) LIGO-Virgo 
  observing runs, and two additional BBHs from the O3a run. We then provide joint constraints with published results
  from the O3a run. In the most agnostic and conservative scenario where we combine the information from different
events using a hierarchical approach, we obtain, at $90\%$ credibility, that 
the fractional deviations in the frequency (damping time) of the dominant QNM are 
$\df{220}=0.02^{+0.09}_{-0.09}$ ($\dtau{220}=0.13^{+0.42}_{-0.40}$), respectively, an improvement of a factor of $\sim 4$ ($\sim 2$) 
over the results obtained with our model in the LIGO-Virgo publication. The single-event most-stringent constraint to date continues to be 
GW150914 for which we obtain $\df{220}=0.04^{+0.06}_{-0.04}$ and $\dtau{220}=0.09^{+0.18}_{-0.18}$.

### 2021-03-25 papers/intro_paper/intro.tex
minor changes to abstract and intro

The consistency between the post-merger signal and the least damped QNM was first demonstrated in
Ref.~\cite{TheLIGOScientific:2016src} for GW150914, and confirmed through several independent analyses~\cite{Brito:2018rfr,Carullo:2019flw,Isi:2019aib,CalderonBustillo:2020tjf}. This was later extended to include
overtones in Refs.~\cite{Giesler:2019uxc,Isi:2019aib,Bhagwat:2019dtm,Forteza:2020hbw,Abbott:2020jks}. Consistency of the late-time waveform with a single QNM is a test of the ringdown of
a BBH coalescence, but not necessarily a test of the no-hair
conjecture, which requires the measurement of (at least) two QNMs (BH
spectroscopy), and consistency between them~\cite{Dreyer:2003bv,Carullo:2018sfu,Carullo:2018gah,Bhagwat:2019bwv,Cabero:2019zyt,Maselli:2019mjd}. The
nature of the remnant object has also been explored through tests of
BH thermodynamics, like the Hawking's area
theorem~\cite{Cabero:2017avf,Isi:2020tac} or through search for echos in the
post-merger signal~\cite{Nielsen:2018lkf,Tsang:2019zra,Lo:2018sep,Abedi:2018npz,Abedi:2020sgg,Testa:2018bzd}.  None of these tests have found evidence for non-BH nature of the remnant
object (as described in GR) in LIGO-Virgo BBH observations.

Ground-based detectors, such as LIGO and Virgo, are most sensitive to
stellar-mass BH binaries that merge near the
minimum of their sensitivity band ($\sim 100$ Hz). As a consequence, the
remnant object \textit{rings down} at frequencies where the sensitivity is worse, leaving
very little signal-to-noise ratio (SNR) in the
post-merger signal. Most of the tests mentioned above restrict the data to this post-merger signal, which signficantly reduces parameter-estimation capabilities. However, added to a lack of SNR is the ambiguity in clearly defining a start time for the ringdown phase (discussed, for example, in Refs.~\cite{Berti:2007fi,Baibhav:2017jhs,Bhagwat:2017tkm}). In some tests~\cite{Carullo:2018gah,Carullo:2019flw}, the ringdown start time has been left as a
free parameter to be estimated directly from the data. Whereas, in other cases, the ringdown start time is predicted from corresponding full-signal parameter-estimation analyses (see \texttt{PyRING} analysis in Ref.~\cite{Abbott:2020jks}). Uncertainties
in estimates of the ringdown start-time, as well as an overall lack of
SNR in the post-merger signal, given typical sensitivities of
ground-based detectors, result in significant statistical uncertainties
in the measurement of the QNM frequencies. And hence, one might want to look at alternate methods to measure QNMs using information from as much as the signal as possible.

>>>>

The consistency between the post-merger signal and the least damped QNM was first demonstrated in
Ref.~\cite{TheLIGOScientific:2016src} for GW150914, and confirmed through several independent analyses~\cite{Brito:2018rfr,Carullo:2019flw,Isi:2019aib,CalderonBustillo:2020tjf}. This was later extended to include
overtones in Refs.~\cite{Giesler:2019uxc,Isi:2019aib,Bhagwat:2019dtm,Forteza:2020hbw,Abbott:2020jks}. Consistency of the late-time waveform with a single QNM is a test of the ringdown of
a BBH coalescence, but not necessarily a test of the no-hair
conjecture, which requires the measurement of (at least) two QNMs (BH
spectroscopy), and consistency between them~\cite{Dreyer:2003bv,Carullo:2018sfu,Carullo:2018gah,Isi:2019aib,Bhagwat:2019bwv,Cabero:2019zyt,Maselli:2019mjd}. The nature of the remnant object has also been explored through tests of BH thermodynamics, like the Hawking's area
theorem~\cite{Cabero:2017avf,Isi:2020tac} \rb{and the Bekenstein-Hod universal bound~\cite{Carullo:2021yxh}, by looking for possible signatures from the BH area quantization~\cite{Foit:2016uxn,Laghi:2020rgl}}, or through search for echos in the
post-merger signal~\cite{Nielsen:2018lkf,Tsang:2019zra,Lo:2018sep,Abedi:2018npz,Abedi:2020sgg,Testa:2018bzd}.  None of these tests have found evidence for non-BH nature of the remnant
object (as described in GR) in LIGO-Virgo BBH observations.

Ground-based detectors, such as LIGO and Virgo, are most sensitive to
stellar-mass BH binaries that merge near the
minimum of their sensitivity band ($\sim 100$ Hz). As a consequence, the
remnant object \textit{rings down} at frequencies where the sensitivity is worse, leaving
very little signal-to-noise ratio (SNR) in the
post-merger signal. Most of the tests mentioned above restrict the data to this post-merger signal, which significantly reduces parameter-estimation capabilities. However, added to a lack of SNR is the ambiguity in clearly defining a start time for the ringdown phase (discussed, for example, in Refs.~\cite{Berti:2007fi,Baibhav:2017jhs,Bhagwat:2017tkm}). In some tests~\cite{Carullo:2018gah,Carullo:2019flw}, the ringdown start time has been left as a
free parameter to be estimated directly from the data. Whereas, in other cases, the ringdown start time is predicted from corresponding full-signal parameter-estimation analyses (see \texttt{PyRING} analysis in Ref.~\cite{Abbott:2020jks}). Uncertainties
in estimates of the ringdown start-time, as well as an overall lack of
SNR in the post-merger signal, given typical sensitivities of
ground-based detectors, result in significant statistical uncertainties
in the measurement of the QNM frequencies. And hence, one might want to look at alternate methods to measure QNMs using information from as much as the signal as possible.
